---
title: "Association Between Sedentary Behavior and Metabolic Dysfunction-Associated Steatotic Liver Disease in U.S. Adults: A Cross-Sectional Analysis of NHANES 2017–2020"
subtitle: "Regression models"
author: "Guilherme Ferreira"
date: "2026-09-12 (YYYY-MM-DD)"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview

This document is part of the Association Between Sedentary Behavior and Metabolic Dysfunction-Associated Steatotic Liver Disease in U.S. Adults: A Cross-Sectional Analysis of NHANES 2017–2020 analysis pipeline and is not intended to be executed independently. It relies on data, derived variables, and objects generated in previous stages of the pipeline.

---

### Variable Naming Convention

* **UPPERCASE variables:** Retain their original NHANES variable names (with missing value codes converted to `NA` and skip patterns handled appropriately).
* **lowercase variables:** Represent derived variables created during processing (e.g., `smoker`).

---

### Key Resources & Documentation

* [NHANES Continuous Data Files (CDC)](https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx)
* [The Epidemiologist R Handbook](https://www.epirhandbook.com/en/)
* [CDC NHANES R Tutorial Examples](https://wwwn.cdc.gov/nchs/data/tutorials/module3_examples_R.r)
* [STC Library Reference](https://libstc.cc/#/stc/nid:6e7ka238xj6qcuyf7dkdkuyfo)
* [R for Medical Public Health Data Analysis](https://bookdown.org/rwnahhas/RMPH/intro.html)


# 1. Loading and Installing Required Packages
```{r Download Packages}

# Clear workspace
rm(list = ls())

# Define required packages
required_packages <- c("survey", "tidyverse", "gtsummary", "gt", "here", "rio", "broom", "car", "conflicted")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
  }
}

# Load all packages silently
suppressPackageStartupMessages({
  invisible(lapply(required_packages, library, character.only = TRUE))
})

# Display current R version
cat("R Version:", as.character(getRversion()), "\n")

```

```{r load-libraries, message=FALSE, warning=FALSE}

# Load required libraries
library(survey)
library(tidyverse)
library(gtsummary)
library(gt)
library(rio)
library(here)
library(broom)
library(car)

library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")

# Display loaded package versions
cat("R package versions:\n")
for (p in required_packages) {
  cat(p, ": ", as.character(packageVersion(p)), "\n")
}

# Configure survey package options
# Specify how the survey package should handle strata containing only one
# Primary Sampling Unit (PSU).

options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)
```


```{r loading-design-and-drop-unused-levels, message=FALSE, warning=FALSE}

load("df_weighted.RData")

# Remove níveis sem observações
svy_design_subset$variables <- svy_design_subset$variables %>%
  mutate(
    alcohol_use = fct_drop(alcohol_use) 
  )
```


```{r Initial-configurations, message=FALSE, warning=FALSE}

# Define dependent variable
dependent_variable <- "masld_288"

descriptive_vars <- c(
  
  # Demographics
  "RIDAGEYR", "RIAGENDR", "race_3", "education", "INDFMPIR",
  
  # Anthropometric
  "BMXBMI", "BMXWAIST", "overweight_obese",
  
  # Blood pressure
  "sbp", "dbp", "htn",
  
  # Lipids
  "LBXTC", "LBDHDD", "LBDLDL", "LBXTR", "low_hdl", "dyslipidemia",
  
  # Glucose
  "LBXGLU", "LBXGH", "dysglycaemia_dm2",
  
  # Lifestyle
  "kcal_mean", "tsugar_adj", "tsatfat_adj", "smoking", "alcohol_use",
  
  # Physical activity
  "mvpa_category", "PAD680", "pad680_hours",
  
  # Liver outcomes
  "LUXCAPM", "LUXSMED", "masld_288", "hepatic_fibrosis"
)

# Define explanatory variables by removing the dependent variable 
# and liver-related variables not included as explanatory variables
explanatory_vars <- setdiff(
  descriptive_vars,
  c(dependent_variable, "hepatic_fibrosis", "LUXSMED", "LUXCAPM", "PAD680")
)

# Create a named vector with custom labels
custom_labels <- list(
  
  # Demographic
  RIDAGEYR = "Age (years)",
  RIAGENDR = "Sex, Male",
  race_3 = "Race/Ethnicity",
  education = "educational level",
  INDFMPIR = "Income-to-Poverty Ratio",
  
  # Anthropometric
  BMXBMI = "BMI (kg/m²)",
  BMXWAIST = "Waist circumference (cm)",
  overweight_obese = "Overweight/Obesity",
  
  # Blood pressure
  sbp = "SBP (mmHg)",
  dbp = "DBP (mmHg)",
  htn = "Hypertension",
  
  # Lipids
  LBXTC = "Total cholesterol (mg/dL)",
  LBDLDL = "LDL-C (mg/dL)",
  LBDHDD = "HDL-C (mg/dL)",
  LBXTR = "Triglycerides (mg/dL)",
  low_hdl = "Low HDL-C",
  dyslipidemia = "dyslipidemia",
  
  # Glucose
  LBXGLU = "Fasting glucose (mg/dL)",
  LBXGH = "HbA1c (%)",
  dysglycaemia_dm2 = "Prediabetes/Diabetes",
  
  # Lifestyle
  kcal_mean = "Total energy intake (kcal/day)",
  tsugar_adj = "Sugar intake",
  tsatfat_adj = "Saturated fat intake",
  smoking = "smoking",
  alcohol_use = "Alcohol use",
  
  # Physical activity
  mvpa_category = "Physical activity (≥150 min/week)",
  PAD680 = "Sedentary behavior (min/day)",
  pad680_hours = "Sedentary behavior (hours/day)",
  
  # Liver outcomes
  LUXCAPM = "CAP (dB/m)",
  LUXSMED = "Liver stiffness (kPa)",
  masld_288 = "MASLD",
  hepatic_fibrosis = "Advanced liver fibrosis"
  
)

```

## 2. Regression Models

### 2.1 Simple Regression Models

```{r map-model-outliers, message=FALSE, warning=FALSE}

# Map influential observations back to the survey design using DFBETAs

adjust_without_outliers <- function(model, design, response, id_var, model_name) {
  
  # Validate model
  if (is.null(model) || !inherits(model, "svyglm")) {
    stop("Model must be a valid svyglm object")
  }
  
  # Calculate weighted DFBETAs
  inf <- influence(model)
  dfbeta_vals <- inf$coefficients
  # Matrix: rows = observations, columns = coefficients
  
  # Cutoff criterion: |DFBETA| > 2/sqrt(n)
  n <- nrow(model.frame(model))
  cutoff <- 2 / sqrt(n)
  
  # Identify influential observations for any coefficient
  influential_cases <- which(
    apply(abs(dfbeta_vals), 1, function(x) any(x > cutoff))
  )
  
  # Plot DFBETAs for the main exposure
  if ("PAD680" %in% colnames(dfbeta_vals)) {
    plot(
      dfbeta_vals[, "PAD680"],
      type = "h",
      main = paste("DFBETA - Model", model_name),
      ylab = "DFBETA for Sedentary Behavior"
    )
    abline(h = c(-cutoff, cutoff), col = "red", lty = 2)
  }
  
  # Obtain model variables
  model_vars <- all.vars(formula(model))
  
  # Map IDs
  model_data <- model.frame(model)
  complete_survey_data <- design$variables[
    complete.cases(design$variables[model_vars]),
  ]
  
  # Obtain IDs of influential observations
  if (length(influential_cases) > 0) {
    outlier_ids <- complete_survey_data[[id_var]][influential_cases]
  } else {
    outlier_ids <- numeric(0)
  }
  
  # Exclude influential observations and refit model
  if (length(outlier_ids) > 0) {
    keep_cases <- !(design$variables[[id_var]] %in% outlier_ids)
    design_without_outliers <- design[keep_cases, ]
  } else {
    design_without_outliers <- design
    message("No influential observations found; using original design")
  }
  
  # Refit model without influential observations
  if (length(outlier_ids) > 0) {
    model_without_outlier <- svyglm(
      formula = formula(model),
      design = design_without_outliers,
      family = model$family
    )
  } else {
    model_without_outlier <- model
  }
  
  # Print information
  print(paste("=== MODEL", model_name, "==="))
  print(paste("Observations in original model:", nrow(model_data)))
  print(paste("Complete survey observations:", nrow(complete_survey_data)))
  print(paste("Influential observations (DFBETA):", length(outlier_ids)))
  
  return(list(
    model_without_outlier = model_without_outlier,
    outliers = outlier_ids,
    cutoff = cutoff,
    influential_cases = influential_cases,
    dfbeta_vals = dfbeta_vals
  ))
}
```
_Note:_ Outlier analysis was performed as a sensitivity analysis. DFBETA-based influence diagnostics are not fully validated for complex survey designs and should therefore not be interpreted as a formal outlier detection procedure for survey-weighted regression models. The results are used to assess the robustness of the primary estimates to potentially influential observations.

```{r simple-regresison-with-and-without-outliers, message=FALSE, warning=FALSE}

# Function to run univariate regression with and without influential observations
# Requires the adjust_without_outliers function

run_univariate <- function(var, design, response, id_var, labels) {
  f <- as.formula(paste(response, "~", var))
  
  # Original model
  model <- svyglm(f, design = design, family = quasibinomial())
  
  # Refit without influential observations
  res <- adjust_without_outliers(
    model, design, response, id_var, var
  )
  
  # Formatted tables
  tbl_original <- tbl_regression(
    model,
    exponentiate = TRUE,
    conf.int = TRUE,
    label = labels
  ) %>%
    add_glance_table(include = c(nobs))
  
  tbl_without_outlier <- tbl_regression(
    res$model_without_outlier,
    exponentiate = TRUE,
    conf.int = TRUE,
    label = labels
  ) %>%
    add_glance_table(include = c(nobs))
  
  return(list(
    original = tbl_original,
    without_outlier = tbl_without_outlier
  ))
}

# Run univariate models for all explanatory variables
univariate_results <- lapply(explanatory_vars, function(v) {
  run_univariate(
    v,
    svy_design_subset,
    dependent_variable,
    "SEQN",
    custom_labels
  )
})

# Stack all tables WITH influential observations
univ_tbl <- purrr::map(
  univariate_results,
  "original"
)

univ_tbl <- tbl_stack(univ_tbl) %>%
  modify_header(
    label ~ "**Variable**",
    estimate ~ "**OR (95% CI)**",
    p.value ~ "**P-value**"
  ) %>%
  modify_caption("**Univariate Analysis - WITH influential observations**") %>%
  bold_labels()

# Stack all tables WITHOUT influential observations
univ_tbl_without_outlier <- purrr::map(
  univariate_results,
  "without_outlier"
)

univ_tbl_without_outlier <- tbl_stack(univ_tbl_without_outlier) %>%
  modify_header(
    label ~ "**Variable**",
    estimate ~ "**OR (95% CI)**",
    p.value ~ "**P-value**"
  ) %>%
  modify_caption("**Univariate Analysis - WITHOUT influential observations**") %>%
  bold_labels()

univ_tbl
univ_tbl_without_outlier

# Export results
export(
  as_tibble(univ_tbl),
  here("Tables", paste0(dependent_variable, "_univariate_table.csv"))
)

export(
  as_tibble(univ_tbl_without_outlier),
  here("Tables", paste0(
    dependent_variable,
    "_univariate_table_without_outlier.csv"
  ))
)

gtsave(
  as_gt(univ_tbl),
  here("Tables", paste0(dependent_variable, "_univariate_table.docx"))
)

gtsave(
  as_gt(univ_tbl),
  here("Tables", paste0(dependent_variable, "_univariate_table.pdf"))
)

gtsave(
  as_gt(univ_tbl_without_outlier),
  here("Tables", paste0(
    dependent_variable,
    "_univariate_table_without_outlier.docx"
  ))
)

gtsave(
  as_gt(univ_tbl_without_outlier),
  here("Tables", paste0(
    dependent_variable,
    "_univariate_table_without_outlier.pdf"
  ))
)
```

## 2.2 Modelos de Regressão Multivariada - Abordagem Aditiva Hierárquica

## 2.2 Multivariable Regression Models - Hierarchical Additive Approach

```{r hierarchical-additive-regression-models, message=FALSE, warning=FALSE}

# Sociodemographic confounders
confounders_model_1 <- c(
  "RIDAGEYR",
  "RIAGENDR",
  "race_3",
  "education",
  "INDFMPIR",
  "smoking",
  "alcohol_use",
  "kcal_mean"
)

# Model 1 + physical activity
confounders_model_2 <- c(
  confounders_model_1,
  "mvpa_category"
)

# Model 2 + metabolic mediators
confounders_model_3 <- c(
  confounders_model_2,
  "overweight_obese",
  "htn",
  "dysglycaemia_dm2",
  "dyslipidemia"
)

# Complete models for regression analyses
model_1_vars <- c("pad680_hours", confounders_model_1)
model_2_vars <- c("pad680_hours", confounders_model_2)
model_3_vars <- c("pad680_hours", confounders_model_3)

model_vars <- list(
  model_1_vars,
  model_2_vars,
  model_3_vars
)

# ====================
# ORIGINAL MODELS
# ====================

regression_models <- list()

for (i in seq_along(model_vars)) {
  vars <- model_vars[[i]]
  formula <- reformulate(
    termlabels = vars,
    response = dependent_variable
  )
  
  regression_models[[i]] <- svyglm(
    formula = formula,
    design = svy_design_subset,
    family = quasibinomial(link = "logit")
  )
}

# Tables for original models
original_tables <- lapply(seq_along(regression_models), function(i) {
  model <- regression_models[[i]]
  n_obs <- nrow(model$model)
  
  tbl_regression(
    model,
    exponentiate = TRUE,
    conf.int = TRUE,
    label = custom_labels
  ) %>%
    modify_caption(
      paste0("**Original Model ", i, " (N = ", n_obs, ")**")
    )
})

# Display original models
tbl_merge(
  tbls = original_tables,
  tab_spanner = paste("Original Model", seq_along(original_tables))
)

# ====================
# MODELS WITHOUT INFLUENTIAL OBSERVATIONS
# ====================

regression_models_without_outliers <- list()

for (i in seq_along(regression_models)) {
  cat("\n=== Processing Model", i, "===\n")
  
  regression_models_without_outliers[[i]] <- adjust_without_outliers(
    model = regression_models[[i]],
    design = svy_design_subset,
    response = dependent_variable,
    id_var = "SEQN",
    model_name = paste0("Total_", i)
  )
}

# Extract refitted models
regression_models_refit <- lapply(
  regression_models_without_outliers,
  function(x) x$model_without_outlier
)

# Create tables for models without influential observations
tables_without_outliers <- lapply(
  seq_along(regression_models_refit),
  function(i) {
    model <- regression_models_refit[[i]]
    
    # Extract number of observations manually
    n_obs <- nrow(model$model)
    
    tbl_regression(
      model,
      exponentiate = TRUE,
      conf.int = TRUE,
      label = custom_labels
    ) %>%
      modify_caption(
        paste0("**Model ", i, " (N = ", n_obs, ")**")
      )
  }
)

# Display models without influential observations
tbl_merge(
  tbls = tables_without_outliers,
  tab_spanner = paste(
    "Model Without Influential Observations",
    seq_along(tables_without_outliers)
  )
)

# ====================
# SIDE-BY-SIDE COMPARISON
# ====================

comparison_table <- tbl_merge(
  tbls = c(original_tables, tables_without_outliers),
  tab_spanner = c(
    paste("Original", seq_along(original_tables)),
    paste("Without Influential Observations",
          seq_along(tables_without_outliers))
  )
)

comparison_table
```


```{r check-removed-outliers, message=FALSE, warning=FALSE}

# Summary of influential observations by model
outlier_summary <- data.frame(
  Model = paste("Model", 1:length(regression_models_without_outliers)),
  N_Outliers = sapply(
    regression_models_without_outliers,
    function(x) length(x$outliers)
  ),
  DFBETA_Cutoff = sapply(
    regression_models_without_outliers,
    function(x) round(x$cutoff, 4)
  )
)

print(outlier_summary)

# Check which IDs were removed from each model
for (i in seq_along(regression_models_without_outliers)) {
  cat("\n=== Model", i, "===\n")
  cat(
    "Removed IDs:",
    if (length(regression_models_without_outliers[[i]]$outliers) > 0) {
      head(
        regression_models_without_outliers[[i]]$outliers,
        10
      )
    } else {
      "None"
    },
    "\n"
  )
}

```

# 2.3 Assessment of Multiple Regression Assumptions

## 2.3.1 Multicollinearity and Linearity in the Logit

```{r regression-diagnostics, message=FALSE, warning=FALSE}

# Regression model diagnostics

# ===== Multicollinearity =====

# Calculate VIF for each model
vif_results <- lapply(regression_models, function(model) {
  tryCatch({
    vif(model)
  }, error = function(e) {
    message("Error calculating VIF: ", e$message)
    return(NULL)
  })
})

# Display results
vif_results

cat(
  "\n=== VIF Interpretation ===\n",
  "VIF ≥ 5 may indicate substantial multicollinearity; ",
  "a more conservative threshold of VIF ≥ 2.5 may also be considered.\n"
)

# ===== Linearity between continuous predictors and the logit =====

crPlots(regression_models[[1]])
crPlots(regression_models[[2]])
crPlots(regression_models[[3]])

cat(
  "\n=== Interpretation of the linearity between continuous predictors and the logit ===\n",
  "\n",
  "- For categorical variables: differences in residual patterns between categories do not represent an assessment of linearity.\n",
  "- For continuous variables: assess whether the smoothed residual trend is approximately consistent with the expected linear relationship.\n"
)
```
**Note:** Multicollinearity and linearity diagnostics were evaluated as exploratory assessments. Standard diagnostic procedures implemented in the `car` package are not specifically validated for complex survey-weighted regression models and should therefore be interpreted cautiously.

## 2.3.2 Linearity Assessment in Multiple Regression

```{r multivariable-regression-linearity-assessment, message=FALSE, warning=FALSE}

# ====================
# Linearity Assessment (quadratic term) - Model 3
# ====================

# Assess whether the relationship between sedentary time (pad680_hours)
# and the logit of MASLD is linear by adding a quadratic term to the
# fully adjusted model (Model 3) and testing its significance using
# a Wald test.

model_3_quadratic <- svyglm(
  reformulate(
    c("pad680_hours", "I(pad680_hours^2)", confounders_model_3),
    dependent_variable
  ),
  design = svy_design_subset,
  family = quasibinomial(link = "logit")
)

p_nonlinearity <- regTermTest(
  model_3_quadratic,
  "I(pad680_hours^2)"
)$p

cat(
  "\n=== Linearity Assessment (quadratic term) - Model 3 ===\n",
  "P-value for the quadratic term:", signif(p_nonlinearity, 3), "\n",
  ifelse(
    p_nonlinearity < 0.05,
    "-> Evidence of nonlinearity (p < 0.05)\n",
    "-> No evidence of nonlinearity (p >= 0.05) - linear relationship is adequate\n"
  )
)
```

```{r predicted-probability-masld, message=FALSE, warning=FALSE}

# ====================
# Figure: Predicted probability of MASLD by sedentary time (Model 3)
# 1st to 99th percentile (P1-P99)
# ====================

model_3_prediction <- regression_models[[3]]

model_data <- svy_design_subset$variables

# --------------------------------------------------------
# Grid of pad680_hours values (1st to 99th percentile)
# --------------------------------------------------------

pad680_interval <- quantile(
  model_data$pad680_hours,
  probs = c(0.01, 0.99),
  na.rm = TRUE
)

pad680_grid <- seq(
  pad680_interval[1],
  pad680_interval[2],
  length.out = 100
)

# --------------------------------------------------------
# Fix remaining covariates at representative values
# --------------------------------------------------------

fix_covariate <- function(x) {
  if (is.numeric(x)) {
    mean(x, na.rm = TRUE)
  } else {
    mode <- names(sort(table(x), decreasing = TRUE))[1]
    factor(mode, levels = levels(as.factor(x)))
  }
}

fixed_covariates <- lapply(
  model_data[confounders_model_3],
  fix_covariate
)

prediction_data <- data.frame(
  pad680_hours = pad680_grid,
  fixed_covariates
)

# --------------------------------------------------------
# Prediction on the logit scale
# --------------------------------------------------------

prediction <- predict(
  model_3_prediction,
  newdata = prediction_data,
  type = "link"
)

# Model matrix for the prediction data
X_prediction <- model.matrix(
  delete.response(terms(model_3_prediction)),
  data = prediction_data
)

# Variance-covariance matrix of model coefficients
vcov_model <- vcov(model_3_prediction)

# Standard errors of the predicted linear predictor
prediction_se <- sqrt(
  rowSums(
    (X_prediction %*% vcov_model) * X_prediction
  )
)

prediction_data <- prediction_data %>%
  mutate(
    logit = as.numeric(prediction),
    standard_error = prediction_se,
    probability = plogis(logit),
    lower_ci = plogis(logit - 1.96 * standard_error),
    upper_ci = plogis(logit + 1.96 * standard_error)
  )

# --------------------------------------------------------
# Plot, with the X-axis explicitly limited to P1-P99
# --------------------------------------------------------

dose_response_plot <- ggplot(
  prediction_data,
  aes(x = pad680_hours, y = probability)
) +
  geom_ribbon(
    aes(ymin = lower_ci, ymax = upper_ci),
    fill = "steelblue",
    alpha = 0.2
  ) +
  geom_line(
    color = "steelblue",
    linewidth = 1
  ) +
  geom_rug(
    data = model_data,
    aes(x = pad680_hours, y = NULL),
    sides = "b",
    alpha = 0.15,
    inherit.aes = FALSE
  ) +
  coord_cartesian(
    xlim = c(pad680_interval[1], pad680_interval[2])
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    x = "Sedentary time (hours/day)",
    y = "Predicted probability of MASLD"
  ) +
  theme_minimal(base_size = 13)

dose_response_plot

# Save figure
ggsave(
  here(
    "Figures",
    "Figure2_dose_response_sedentary_behavior_MASLD.png"
  ),
  plot = dose_response_plot,
  width = 7,
  height = 5,
  dpi = 300
)

extreme_predictions <- prediction_data %>%
  slice(c(1, n())) %>%
  select(
    pad680_hours,
    probability,
    lower_ci,
    upper_ci
  )

print(extreme_predictions)

cat(
  sprintf(
    "Sedentary time = %.1f h/day: predicted probability = %.1f%% (95%% CI: %.1f-%.1f%%)\n",
    extreme_predictions$pad680_hours[1],
    extreme_predictions$probability[1] * 100,
    extreme_predictions$lower_ci[1] * 100,
    extreme_predictions$upper_ci[1] * 100
  ),
  sprintf(
    "Sedentary time = %.1f h/day: predicted probability = %.1f%% (95%% CI: %.1f-%.1f%%)\n",
    extreme_predictions$pad680_hours[2],
    extreme_predictions$probability[2] * 100,
    extreme_predictions$lower_ci[2] * 100,
    extreme_predictions$upper_ci[2] * 100
  )
)
```

# 2.4 Análise exploratória - Subgrupos

```{r exploratory-subgroup-analysis, message=FALSE, warning=FALSE}}

# ============================================================
# Custom labels specific to this analysis
# ============================================================

custom_labels_2 <- c(
  custom_labels,
  list(
    ridageyr_cat  = "Age Group",
    indfmpir_cat = "Poverty Income Ratio (category)",
    kcal_mean_cat = "Total Energy Intake (category)"
  )
)


# ============================================================
# Helper function to format p-values
# ============================================================

format_p <- function(p, decimals = 3) {
  
  ifelse(
    is.na(p),
    NA_character_,
    ifelse(
      p < 10^(-decimals),
      paste0("<0.", strrep("0", decimals - 1), "1"),
      sprintf(paste0("%.", decimals, "f"), p)
    )
  )
}


# ============================================================
# Helper function to remove covariates that are constant
# within a subgroup
# ============================================================

remove_constant_covariates <- function(design_sub, variables) {
  
  keep <- sapply(variables, function(v) {
    
    values <- na.omit(design_sub$variables[[v]])
    
    if (is.factor(values) || is.character(values)) {
      length(unique(values)) >= 2
    } else {
      TRUE
    }
    
  })
  
  removed <- variables[!keep]
  
  if (length(removed) > 0) {
    cat(
      "    Removing covariate(s) that are constant within the subgroup:",
      paste(removed, collapse = ", "),
      "\n"
    )
  }
  
  variables[keep]
}


# ============================================================
# Pairs of categorical and continuous variables to exclude
# simultaneously
# ============================================================

variable_pairs <- list(
  ridageyr_cat  = c("ridageyr_cat", "RIDAGEYR"),
  kcal_mean_cat = c("kcal_mean_cat", "kcal_mean"),
  indfmpir_cat = c("indfmpir_cat", "INDFMPIR")
)


# ============================================================
# Subgroup variables
# ============================================================

subgroups <- c(
  "mvpa_category",
  "RIAGENDR",
  "ridageyr_cat",
  "race_3",
  "education",
  "indfmpir_cat",
  "overweight_obese",
  "dysglycaemia_dm2",
  "htn",
  "dyslipidemia",
  "kcal_mean_cat",
  "smoking",
  "alcohol_use"
)


# ============================================================
# Main function
# ============================================================

subgroup_analysis <- function(
    design,
    response,
    exposure = "pad680_hours",
    subgroups,
    adjustment
) {
  
  final_results <- list()
  
  
  for (subgroup in subgroups) {
    
    cat("\n=============================\n")
    cat("Subgroup:", subgroup, "\n")
    
    
    adjustment_interaction <- setdiff(
      adjustment,
      if (!is.null(variable_pairs[[subgroup]])) {
        variable_pairs[[subgroup]]
      } else {
        subgroup
      }
    )
    
    
    # --------------------------------------------------------
    # Interaction model
    # --------------------------------------------------------
    
    p_interaction <- tryCatch({
      
      interaction_formula <- reformulate(
        c(
          exposure,
          subgroup,
          adjustment_interaction,
          paste0(exposure, ":", subgroup)
        ),
        response = response
      )
      
      interaction_model <- svyglm(
        interaction_formula,
        design = design,
        family = quasibinomial()
      )
      
      regTermTest(
        interaction_model,
        paste0(exposure, ":", subgroup)
      )$p
      
    }, error = function(e) {
      
      cat(
        "Error in the interaction model for subgroup",
        subgroup,
        ":",
        e$message,
        "\n"
      )
      
      NA
    })
    
    
    # --------------------------------------------------------
    # Identify valid subgroup levels
    # --------------------------------------------------------
    
    levels <- unique(
      na.omit(
        design$variables[[subgroup]]
      )
    )
    
    
    for (level in levels) {
      
      cat("  Category:", level, "\n")
      
      
      tryCatch({
        
        
        # ----------------------------------------------------
        # Create subgroup-specific survey design
        # ----------------------------------------------------
        
        indices <- which(
          !is.na(design$variables[[subgroup]]) &
            design$variables[[subgroup]] == level
        )
        
        
        if (length(indices) == 0)
          next
        
        
        design_sub <- design[indices, ]
        
        
        # Check survey weights
        if (any(is.na(weights(design_sub)))) {
          cat("NA survey weight found - skipping category\n")
          next
        }
        
        
        # Minimum sample size
        if (nrow(design_sub$variables) < 50)
          next
        
        
        # ----------------------------------------------------
        # Remove covariates that are constant within subgroup
        # ----------------------------------------------------
        
        adjustment_sub <- remove_constant_covariates(
          design_sub,
          adjustment_interaction
        )
        
        
        # ----------------------------------------------------
        # Stratified model
        # ----------------------------------------------------
        
        subgroup_formula <- reformulate(
          c(
            exposure,
            adjustment_sub
          ),
          response = response
        )
        
        
        subgroup_model <- svyglm(
          subgroup_formula,
          design = design_sub,
          family = quasibinomial()
        )
        
        
        # ----------------------------------------------------
        # Extract exposure effect
        # ----------------------------------------------------
        
        result <- broom::tidy(
          subgroup_model,
          exponentiate = TRUE,
          conf.int = TRUE
        ) %>%
          filter(term == exposure) %>%
          mutate(
            
            Subgroup = subgroup,
            
            Category = level,
            
            N = nobs(subgroup_model),
            
            OR_CI95 = sprintf(
              "%.2f (%.2f–%.2f)",
              estimate,
              conf.low,
              conf.high
            ),
            
            P = format_p(p.value),
            
            P_interaction = ifelse(
              level == levels[1],
              format_p(p_interaction),
              NA_character_
            )
            
          ) %>%
          select(
            Subgroup,
            Category,
            N,
            OR_CI95,
            P,
            P_interaction
          )
        
        
        final_results[[paste0(subgroup, "_", level)]] <- result
        
        
      }, error = function(e) {
        
        cat(
          "Error in category",
          level,
          "of subgroup",
          subgroup,
          ":",
          e$message,
          "\n"
        )
        
      })
      
      
    }
    
    
  }
  
  table <- bind_rows(final_results)
  
  # ----------------------------------------------------------
  # Apply readable labels to subgroup names
  # ----------------------------------------------------------
  
  labels_vec <- unlist(custom_labels_2)
  
  table$Subgroup <- ifelse(
    !is.na(labels_vec[table$Subgroup]),
    labels_vec[table$Subgroup],
    table$Subgroup
  )
  
  table
}


# ============================================================
# Run subgroup analysis
# ============================================================

subgroup_table <- subgroup_analysis(
  
  design = svy_design_subset,
  
  response = dependent_variable,
  
  exposure = "pad680_hours",
  
  subgroups = subgroups,
  
  adjustment = confounders_model_3
)


# Display results

subgroup_table
```