---
title: "Association Between Sedentary Behavior and Metabolic Dysfunction-Associated Steatotic Liver Disease in U.S. Adults: A Cross-Sectional Analysis of NHANES 2017–2020"
subtitle: "Sensitivity analysis"
author: "Guilherme Ferreira"
date: "2026-09-12 (YYYY-MM-DD)"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview

This document is part of the Association Between Sedentary Behavior and Metabolic Dysfunction-Associated Steatotic Liver Disease in U.S. Adults: A Cross-Sectional Analysis of NHANES 2017–2020 analysis pipeline and is not intended to be executed independently. It relies on data, derived variables, and objects generated in previous stages of the pipeline.

---

### Variable Naming Convention

* **UPPERCASE variables:** Retain their original NHANES variable names (with missing value codes converted to `NA` and skip patterns handled appropriately).
* **lowercase variables:** Represent derived variables created during processing (e.g., `smoker`).

---

### Key Resources & Documentation

* [NHANES Continuous Data Files (CDC)](https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx)
* [The Epidemiologist R Handbook](https://www.epirhandbook.com/en/)
* [CDC NHANES R Tutorial Examples](https://wwwn.cdc.gov/nchs/data/tutorials/module3_examples_R.r)
* [STC Library Reference](https://libstc.cc/#/stc/nid:6e7ka238xj6qcuyf7dkdkuyfo)
* [R for Medical Public Health Data Analysis](https://bookdown.org/rwnahhas/RMPH/intro.html)


# 1. Loading and Installing Required Packages
```{r Download Packages}

# Clear workspace
rm(list = ls())

# Define required packages
required_packages <- c("survey", "tidyverse", "gtsummary", "gt", "here", "rio", "broom", "car", "conflicted")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
  }
}

# Load all packages silently
suppressPackageStartupMessages({
  invisible(lapply(required_packages, library, character.only = TRUE))
})

# Display current R version
cat("R Version:", as.character(getRversion()), "\n")

```

```{r load-libraries, message=FALSE, warning=FALSE}

# Load required libraries
library(survey)
library(tidyverse)
library(gtsummary)
library(gt)
library(rio)
library(here)
library(broom)
library(car)

library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")

# Display loaded package versions
cat("R package versions:\n")
for (p in required_packages) {
  cat(p, ": ", as.character(packageVersion(p)), "\n")
}

# Configure survey package options
# Specify how the survey package should handle strata containing only one
# Primary Sampling Unit (PSU).

options(survey.lonely.psu = "adjust")
options(survey.adjust.domain.lonely = TRUE)
```


```{r loading-design-and-drop-unused-levels, message=FALSE, warning=FALSE}

load("df_weighted.RData")

# Remove níveis sem observações
svy_design_subset$variables <- svy_design_subset$variables %>%
  mutate(
    alcohol_use = fct_drop(alcohol_use) 
  )
```

2. Cutoff: 248 db/m
```{r Initial-configurations, message=FALSE, warning=FALSE}

# Define dependent variable
dependent_variable <- "masld_248"

descriptive_vars <- c(
  
  # Demographics
  "RIDAGEYR", "RIAGENDR", "race_3", "education", "INDFMPIR",
  
  # Anthropometric
  "BMXBMI", "BMXWAIST", "overweight_obese",
  
  # Blood pressure
  "sbp", "dbp", "htn",
  
  # Lipids
  "LBXTC", "LBDHDD", "LBDLDL", "LBXTR", "low_hdl", "dyslipidemia",
  
  # Glucose
  "LBXGLU", "LBXGH", "dysglycaemia_dm2",
  
  # Lifestyle
  "kcal_mean", "tsugar_adj", "tsatfat_adj", "smoking", "alcohol_use",
  
  # Physical activity
  "mvpa_category", "PAD680", "pad680_hours",
  
  # Liver outcomes
  "LUXCAPM", "LUXSMED", "masld_248", "hepatic_fibrosis"
)

# Define explanatory variables by removing the dependent variable 
# and liver-related variables not included as explanatory variables
explanatory_vars <- setdiff(
  descriptive_vars,
  c(dependent_variable, "hepatic_fibrosis", "LUXSMED", "LUXCAPM", "PAD680")
)

# Create a named vector with custom labels
custom_labels <- list(
  
  # Demographic
  RIDAGEYR = "Age (years)",
  RIAGENDR = "Sex, Male",
  race_3 = "Race/Ethnicity",
  education = "educational level",
  INDFMPIR = "Income-to-Poverty Ratio",
  
  # Anthropometric
  BMXBMI = "BMI (kg/m²)",
  BMXWAIST = "Waist circumference (cm)",
  overweight_obese = "Overweight/Obesity",
  
  # Blood pressure
  sbp = "SBP (mmHg)",
  dbp = "DBP (mmHg)",
  htn = "Hypertension",
  
  # Lipids
  LBXTC = "Total cholesterol (mg/dL)",
  LBDLDL = "LDL-C (mg/dL)",
  LBDHDD = "HDL-C (mg/dL)",
  LBXTR = "Triglycerides (mg/dL)",
  low_hdl = "Low HDL-C",
  dyslipidemia = "dyslipidemia",
  
  # Glucose
  LBXGLU = "Fasting glucose (mg/dL)",
  LBXGH = "HbA1c (%)",
  dysglycaemia_dm2 = "Prediabetes/Diabetes",
  
  # Lifestyle
  kcal_mean = "Total energy intake (kcal/day)",
  tsugar_adj = "Sugar intake",
  tsatfat_adj = "Saturated fat intake",
  smoking = "smoking",
  alcohol_use = "Alcohol use",
  
  # Physical activity
  mvpa_category = "Physical activity (≥150 min/week)",
  PAD680 = "Sedentary behavior (min/day)",
  pad680_hours = "Sedentary behavior (hours/day)",
  
  # Liver outcomes
  LUXCAPM = "CAP (dB/m)",
  LUXSMED = "Liver stiffness (kPa)",
  masld_248 = "MASLD",
  hepatic_fibrosis = "Advanced liver fibrosis"
  
)

```

## 2. Cutoff: 248 - Regression Models

### 2.1 Simple Regression Models

```{r map-model-outliers, message=FALSE, warning=FALSE}

# Map influential observations back to the survey design using DFBETAs

adjust_without_outliers <- function(model, design, response, id_var, model_name) {
  
  # Validate model
  if (is.null(model) || !inherits(model, "svyglm")) {
    stop("Model must be a valid svyglm object")
  }
  
  # Calculate weighted DFBETAs
  inf <- influence(model)
  dfbeta_vals <- inf$coefficients
  # Matrix: rows = observations, columns = coefficients
  
  # Cutoff criterion: |DFBETA| > 2/sqrt(n)
  n <- nrow(model.frame(model))
  cutoff <- 2 / sqrt(n)
  
  # Identify influential observations for any coefficient
  influential_cases <- which(
    apply(abs(dfbeta_vals), 1, function(x) any(x > cutoff))
  )
  
  # Plot DFBETAs for the main exposure
  if ("PAD680" %in% colnames(dfbeta_vals)) {
    plot(
      dfbeta_vals[, "PAD680"],
      type = "h",
      main = paste("DFBETA - Model", model_name),
      ylab = "DFBETA for Sedentary Behavior"
    )
    abline(h = c(-cutoff, cutoff), col = "red", lty = 2)
  }
  
  # Obtain model variables
  model_vars <- all.vars(formula(model))
  
  # Map IDs
  model_data <- model.frame(model)
  complete_survey_data <- design$variables[
    complete.cases(design$variables[model_vars]),
  ]
  
  # Obtain IDs of influential observations
  if (length(influential_cases) > 0) {
    outlier_ids <- complete_survey_data[[id_var]][influential_cases]
  } else {
    outlier_ids <- numeric(0)
  }
  
  # Exclude influential observations and refit model
  if (length(outlier_ids) > 0) {
    keep_cases <- !(design$variables[[id_var]] %in% outlier_ids)
    design_without_outliers <- design[keep_cases, ]
  } else {
    design_without_outliers <- design
    message("No influential observations found; using original design")
  }
  
  # Refit model without influential observations
  if (length(outlier_ids) > 0) {
    model_without_outlier <- svyglm(
      formula = formula(model),
      design = design_without_outliers,
      family = model$family
    )
  } else {
    model_without_outlier <- model
  }
  
  # Print information
  print(paste("=== MODEL", model_name, "==="))
  print(paste("Observations in original model:", nrow(model_data)))
  print(paste("Complete survey observations:", nrow(complete_survey_data)))
  print(paste("Influential observations (DFBETA):", length(outlier_ids)))
  
  return(list(
    model_without_outlier = model_without_outlier,
    outliers = outlier_ids,
    cutoff = cutoff,
    influential_cases = influential_cases,
    dfbeta_vals = dfbeta_vals
  ))
}
```
_Note:_ Outlier analysis was performed as a sensitivity analysis. DFBETA-based influence diagnostics are not fully validated for complex survey designs and should therefore not be interpreted as a formal outlier detection procedure for survey-weighted regression models. The results are used to assess the robustness of the primary estimates to potentially influential observations.

```{r simple-sensitivity-analysis, message=FALSE, warning=FALSE}

# ====================
# Sensitivity analysis: simple regression
# With and without influential observations
# ====================

# Simple regression model
simple_model <- svyglm(
  reformulate(
    "pad680_hours",
    dependent_variable
  ),
  design = svy_design_subset,
  family = quasibinomial()
)

# Refit without influential observations
simple_result <- adjust_without_outliers(
  model = simple_model,
  design = svy_design_subset,
  response = dependent_variable,
  id_var = "SEQN",
  model_name = "Simple Regression"
)

# Results with influential observations
simple_table <- tbl_regression(
  simple_model,
  exponentiate = TRUE,
  conf.int = TRUE,
  label = custom_labels
) %>%
  add_glance_table(include = nobs)

# Results without influential observations
simple_table_without_outliers <- tbl_regression(
  simple_result$model_without_outlier,
  exponentiate = TRUE,
  conf.int = TRUE,
  label = custom_labels
) %>%
  add_glance_table(include = nobs)

# Display results
simple_table
simple_table_without_outliers
```

## 2.2 Cutoff: 248 - Multivariable Regression Models - Hierarchical Additive Approach

```{r sensitivity-hierarchical-additive-regression-models, message=FALSE, warning=FALSE}

# Sociodemographic confounders
sensitivity_confounders_model_1 <- c(
  "RIDAGEYR",
  "RIAGENDR",
  "race_3",
  "education",
  "INDFMPIR",
  "smoking",
  "alcohol_use",
  "kcal_mean"
)

# Model 1 + physical activity
sensitivity_confounders_model_2 <- c(
  sensitivity_confounders_model_1,
  "mvpa_category"
)

# Model 2 + metabolic mediators
sensitivity_confounders_model_3 <- c(
  sensitivity_confounders_model_2,
  "overweight_obese",
  "htn",
  "dysglycaemia_dm2",
  "dyslipidemia"
)

# Complete sensitivity models
sensitivity_model_1_vars <- c(
  "pad680_hours",
  sensitivity_confounders_model_1
)

sensitivity_model_2_vars <- c(
  "pad680_hours",
  sensitivity_confounders_model_2
)

sensitivity_model_3_vars <- c(
  "pad680_hours",
  sensitivity_confounders_model_3
)

sensitivity_model_vars <- list(
  sensitivity_model_1_vars,
  sensitivity_model_2_vars,
  sensitivity_model_3_vars
)

# ====================
# ORIGINAL SENSITIVITY MODELS
# ====================

sensitivity_models <- list()

for (i in seq_along(sensitivity_model_vars)) {
  
  formula <- reformulate(
    termlabels = sensitivity_model_vars[[i]],
    response = dependent_variable
  )
  
  sensitivity_models[[i]] <- svyglm(
    formula = formula,
    design = svy_design_subset,
    family = quasibinomial(link = "logit")
  )
}

# Tables for original sensitivity models
sensitivity_original_tables <- lapply(
  seq_along(sensitivity_models),
  function(i) {
    
    model <- sensitivity_models[[i]]
    n_obs <- nrow(model$model)
    
    tbl_regression(
      model,
      exponentiate = TRUE,
      conf.int = TRUE,
      label = custom_labels
    ) %>%
      modify_caption(
        paste0(
          "**Sensitivity Model ", i,
          " (N = ", n_obs, ")**"
        )
      )
  }
)

# Display original sensitivity models
tbl_merge(
  tbls = sensitivity_original_tables,
  tab_spanner = paste(
    "Sensitivity Model",
    seq_along(sensitivity_original_tables)
  )
)

# ====================
# SENSITIVITY MODELS WITHOUT INFLUENTIAL OBSERVATIONS
# ====================

sensitivity_models_without_outliers <- list()

for (i in seq_along(sensitivity_models)) {
  
  cat("\n=== Processing Sensitivity Model", i, "===\n")
  
  sensitivity_models_without_outliers[[i]] <- adjust_without_outliers(
    model = sensitivity_models[[i]],
    design = svy_design_subset,
    response = dependent_variable,
    id_var = "SEQN",
    model_name = paste0("Sensitivity_", i)
  )
}

# Extract refitted models
sensitivity_models_refit <- lapply(
  sensitivity_models_without_outliers,
  function(x) x$model_without_outlier
)

# Create tables for sensitivity models without influential observations
sensitivity_tables_without_outliers <- lapply(
  seq_along(sensitivity_models_refit),
  function(i) {
    
    model <- sensitivity_models_refit[[i]]
    n_obs <- nrow(model$model)
    
    tbl_regression(
      model,
      exponentiate = TRUE,
      conf.int = TRUE,
      label = custom_labels
    ) %>%
      modify_caption(
        paste0(
          "**Sensitivity Model ", i,
          " Without Influential Observations (N = ",
          n_obs,
          ")**"
        )
      )
  }
)

# Display sensitivity models without influential observations
tbl_merge(
  tbls = sensitivity_tables_without_outliers,
  tab_spanner = paste(
    "Sensitivity Model Without Influential Observations",
    seq_along(sensitivity_tables_without_outliers)
  )
)

# ====================
# SIDE-BY-SIDE COMPARISON
# ====================

sensitivity_comparison_table <- tbl_merge(
  tbls = c(
    sensitivity_original_tables,
    sensitivity_tables_without_outliers
  ),
  tab_spanner = c(
    paste(
      "Original Sensitivity Model",
      seq_along(sensitivity_original_tables)
    ),
    paste(
      "Sensitivity Model Without Influential Observations",
      seq_along(sensitivity_tables_without_outliers)
    )
  )
)

sensitivity_comparison_table
``` 

```{r check-sensitivity-outliers, message=FALSE, warning=FALSE}

# Summary of influential observations by sensitivity model
sensitivity_outlier_summary <- data.frame(
  Model = paste(
    "Sensitivity Model",
    1:length(sensitivity_models_without_outliers)
  ),
  N_Outliers = sapply(
    sensitivity_models_without_outliers,
    function(x) length(x$outliers)
  ),
  DFBETA_Cutoff = sapply(
    sensitivity_models_without_outliers,
    function(x) round(x$cutoff, 4)
  )
)

print(sensitivity_outlier_summary)

# Check which IDs were removed from each sensitivity model
for (i in seq_along(sensitivity_models_without_outliers)) {
  
  cat("\n=== Sensitivity Model", i, "===\n")
  
  cat(
    "Removed IDs:",
    if (length(sensitivity_models_without_outliers[[i]]$outliers) > 0) {
      head(
        sensitivity_models_without_outliers[[i]]$outliers,
        10
      )
    } else {
      "None"
    },
    "\n"
  )
}
```

2. Cutoff: 302 db/m
```{r Initial-configurations-2, message=FALSE, warning=FALSE}

# Define dependent variable
dependent_variable <- "masld_302"

descriptive_vars <- c(
  
  # Demographics
  "RIDAGEYR", "RIAGENDR", "race_3", "education", "INDFMPIR",
  
  # Anthropometric
  "BMXBMI", "BMXWAIST", "overweight_obese",
  
  # Blood pressure
  "sbp", "dbp", "htn",
  
  # Lipids
  "LBXTC", "LBDHDD", "LBDLDL", "LBXTR", "low_hdl", "dyslipidemia",
  
  # Glucose
  "LBXGLU", "LBXGH", "dysglycaemia_dm2",
  
  # Lifestyle
  "kcal_mean", "tsugar_adj", "tsatfat_adj", "smoking", "alcohol_use",
  
  # Physical activity
  "mvpa_category", "PAD680", "pad680_hours",
  
  # Liver outcomes
  "LUXCAPM", "LUXSMED", "masld_302", "hepatic_fibrosis"
)

# Define explanatory variables by removing the dependent variable 
# and liver-related variables not included as explanatory variables
explanatory_vars <- setdiff(
  descriptive_vars,
  c(dependent_variable, "hepatic_fibrosis", "LUXSMED", "LUXCAPM", "PAD680")
)

# Create a named vector with custom labels
custom_labels <- list(
  
  # Demographic
  RIDAGEYR = "Age (years)",
  RIAGENDR = "Sex, Male",
  race_3 = "Race/Ethnicity",
  education = "educational level",
  INDFMPIR = "Income-to-Poverty Ratio",
  
  # Anthropometric
  BMXBMI = "BMI (kg/m²)",
  BMXWAIST = "Waist circumference (cm)",
  overweight_obese = "Overweight/Obesity",
  
  # Blood pressure
  sbp = "SBP (mmHg)",
  dbp = "DBP (mmHg)",
  htn = "Hypertension",
  
  # Lipids
  LBXTC = "Total cholesterol (mg/dL)",
  LBDLDL = "LDL-C (mg/dL)",
  LBDHDD = "HDL-C (mg/dL)",
  LBXTR = "Triglycerides (mg/dL)",
  low_hdl = "Low HDL-C",
  dyslipidemia = "dyslipidemia",
  
  # Glucose
  LBXGLU = "Fasting glucose (mg/dL)",
  LBXGH = "HbA1c (%)",
  dysglycaemia_dm2 = "Prediabetes/Diabetes",
  
  # Lifestyle
  kcal_mean = "Total energy intake (kcal/day)",
  tsugar_adj = "Sugar intake",
  tsatfat_adj = "Saturated fat intake",
  smoking = "smoking",
  alcohol_use = "Alcohol use",
  
  # Physical activity
  mvpa_category = "Physical activity (≥150 min/week)",
  PAD680 = "Sedentary behavior (min/day)",
  pad680_hours = "Sedentary behavior (hours/day)",
  
  # Liver outcomes
  LUXCAPM = "CAP (dB/m)",
  LUXSMED = "Liver stiffness (kPa)",
  masld_302 = "MASLD",
  hepatic_fibrosis = "Advanced liver fibrosis"
  
)

```

## 3. Cutoff: 302 - Regression Models

### 3.1 Simple Regression Models


```{r simple-sensitivity-analysis-2, message=FALSE, warning=FALSE}

# ====================
# Sensitivity analysis: simple regression
# With and without influential observations
# ====================

# Simple regression model
simple_model <- svyglm(
  reformulate(
    "pad680_hours",
    dependent_variable
  ),
  design = svy_design_subset,
  family = quasibinomial()
)

# Refit without influential observations
simple_result <- adjust_without_outliers(
  model = simple_model,
  design = svy_design_subset,
  response = dependent_variable,
  id_var = "SEQN",
  model_name = "Simple Regression"
)

# Results with influential observations
simple_table <- tbl_regression(
  simple_model,
  exponentiate = TRUE,
  conf.int = TRUE,
  label = custom_labels
) %>%
  add_glance_table(include = nobs)

# Results without influential observations
simple_table_without_outliers <- tbl_regression(
  simple_result$model_without_outlier,
  exponentiate = TRUE,
  conf.int = TRUE,
  label = custom_labels
) %>%
  add_glance_table(include = nobs)

# Display results
simple_table
simple_table_without_outliers
```

## 3.2 Cutoff: 302 - Multivariable Regression Models - Hierarchical Additive Approach

```{r sensitivity-hierarchical-additive-regression-models-2, message=FALSE, warning=FALSE}

# Sociodemographic confounders
sensitivity_confounders_model_1 <- c(
  "RIDAGEYR",
  "RIAGENDR",
  "race_3",
  "education",
  "INDFMPIR",
  "smoking",
  "alcohol_use",
  "kcal_mean"
)

# Model 1 + physical activity
sensitivity_confounders_model_2 <- c(
  sensitivity_confounders_model_1,
  "mvpa_category"
)

# Model 2 + metabolic mediators
sensitivity_confounders_model_3 <- c(
  sensitivity_confounders_model_2,
  "overweight_obese",
  "htn",
  "dysglycaemia_dm2",
  "dyslipidemia"
)

# Complete sensitivity models
sensitivity_model_1_vars <- c(
  "pad680_hours",
  sensitivity_confounders_model_1
)

sensitivity_model_2_vars <- c(
  "pad680_hours",
  sensitivity_confounders_model_2
)

sensitivity_model_3_vars <- c(
  "pad680_hours",
  sensitivity_confounders_model_3
)

sensitivity_model_vars <- list(
  sensitivity_model_1_vars,
  sensitivity_model_2_vars,
  sensitivity_model_3_vars
)

# ====================
# ORIGINAL SENSITIVITY MODELS
# ====================

sensitivity_models <- list()

for (i in seq_along(sensitivity_model_vars)) {
  
  formula <- reformulate(
    termlabels = sensitivity_model_vars[[i]],
    response = dependent_variable
  )
  
  sensitivity_models[[i]] <- svyglm(
    formula = formula,
    design = svy_design_subset,
    family = quasibinomial(link = "logit")
  )
}

# Tables for original sensitivity models
sensitivity_original_tables <- lapply(
  seq_along(sensitivity_models),
  function(i) {
    
    model <- sensitivity_models[[i]]
    n_obs <- nrow(model$model)
    
    tbl_regression(
      model,
      exponentiate = TRUE,
      conf.int = TRUE,
      label = custom_labels
    ) %>%
      modify_caption(
        paste0(
          "**Sensitivity Model ", i,
          " (N = ", n_obs, ")**"
        )
      )
  }
)

# Display original sensitivity models
tbl_merge(
  tbls = sensitivity_original_tables,
  tab_spanner = paste(
    "Sensitivity Model",
    seq_along(sensitivity_original_tables)
  )
)

# ====================
# SENSITIVITY MODELS WITHOUT INFLUENTIAL OBSERVATIONS
# ====================

sensitivity_models_without_outliers <- list()

for (i in seq_along(sensitivity_models)) {
  
  cat("\n=== Processing Sensitivity Model", i, "===\n")
  
  sensitivity_models_without_outliers[[i]] <- adjust_without_outliers(
    model = sensitivity_models[[i]],
    design = svy_design_subset,
    response = dependent_variable,
    id_var = "SEQN",
    model_name = paste0("Sensitivity_", i)
  )
}

# Extract refitted models
sensitivity_models_refit <- lapply(
  sensitivity_models_without_outliers,
  function(x) x$model_without_outlier
)

# Create tables for sensitivity models without influential observations
sensitivity_tables_without_outliers <- lapply(
  seq_along(sensitivity_models_refit),
  function(i) {
    
    model <- sensitivity_models_refit[[i]]
    n_obs <- nrow(model$model)
    
    tbl_regression(
      model,
      exponentiate = TRUE,
      conf.int = TRUE,
      label = custom_labels
    ) %>%
      modify_caption(
        paste0(
          "**Sensitivity Model ", i,
          " Without Influential Observations (N = ",
          n_obs,
          ")**"
        )
      )
  }
)

# Display sensitivity models without influential observations
tbl_merge(
  tbls = sensitivity_tables_without_outliers,
  tab_spanner = paste(
    "Sensitivity Model Without Influential Observations",
    seq_along(sensitivity_tables_without_outliers)
  )
)

# ====================
# SIDE-BY-SIDE COMPARISON
# ====================

sensitivity_comparison_table <- tbl_merge(
  tbls = c(
    sensitivity_original_tables,
    sensitivity_tables_without_outliers
  ),
  tab_spanner = c(
    paste(
      "Original Sensitivity Model",
      seq_along(sensitivity_original_tables)
    ),
    paste(
      "Sensitivity Model Without Influential Observations",
      seq_along(sensitivity_tables_without_outliers)
    )
  )
)

sensitivity_comparison_table
``` 

```{r check-sensitivity-outliers-2, message=FALSE, warning=FALSE}

# Summary of influential observations by sensitivity model
sensitivity_outlier_summary <- data.frame(
  Model = paste(
    "Sensitivity Model",
    1:length(sensitivity_models_without_outliers)
  ),
  N_Outliers = sapply(
    sensitivity_models_without_outliers,
    function(x) length(x$outliers)
  ),
  DFBETA_Cutoff = sapply(
    sensitivity_models_without_outliers,
    function(x) round(x$cutoff, 4)
  )
)

print(sensitivity_outlier_summary)

# Check which IDs were removed from each sensitivity model
for (i in seq_along(sensitivity_models_without_outliers)) {
  
  cat("\n=== Sensitivity Model", i, "===\n")
  
  cat(
    "Removed IDs:",
    if (length(sensitivity_models_without_outliers[[i]]$outliers) > 0) {
      head(
        sensitivity_models_without_outliers[[i]]$outliers,
        10
      )
    } else {
      "None"
    },
    "\n"
  )
}
```


